Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-11-16 prop_letter.tex
nov 16

\documentclass[english,reprint, aps, prl,superscriptaddress, showpacs,
showkeys, longbibliography, amsmath, amssymb, floatfix]{revtex4-1} 
\usepackage{fullpage}
\usepackage{cmap}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[french,main=english]{babel}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{bbold}
\usepackage{wesa}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage[backref=false]{hyperref}
\usepackage{booktabs}
\usepackage{multirow}
\DeclareUnicodeCharacter{0229}{\c{e}}

@@

\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\qevents}{\ensuremath{\mathcal{E}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\imposs}{{\text{\wesa{impossible}}}}
\newcommand{\likely}{{\text{\wesa{likely}}}}
\newcommand{\unlikely}{{\text{\wesa{unlikely}}}}
\newcommand{\necess}{{\text{\wesa{certain}}}}
\newcommand{\unknown}{{\text{\wesa{unknown}}}}
\newcommand{\midd}{{\text{\wesa{middle}}}}
\usepackage[braket,qm]{qcircuit}
\newcommand{\fket}[1]{{|#1\rangle}}
\newcommand{\fproj}[1]{|#1\rangle\langle #1|}
\newcommand{\proj}[1]{\op{#1}{#1}}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{framed}
\newcommand{\amr}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{green}{Amr says: #1}\end{minipage}\end{framed}}
\newcommand{\yutsung}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}\end{framed}}
\newcommand{\gerardo}[1]{\fbox{\begin{minipage}{0.9\linewidth}\color{OliveGreen}{Gerardo says: #1}\end{minipage}}}
\newcommand{\andy}[1]{\fbox{\begin{minipage}{0.9\linewidth}\color{blue}{Andy says: #1}\end{minipage}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\allowdisplaybreaks
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{L{#1}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{R{#1}}}}
\setcounter{secnumdepth}{3}
% https://tex.stackexchange.com/questions/345284/can-i-have-an-almost-non-breaking-space-in-latex
% Change non-breaking space to \nolinebreak[1] 
% Hence, it will not hyphenate too much surrounding words.
\newcommand{\rmd}{d}
\newcommand{\rmi}{i}
\lineskiplimit=1pt
\newcommand{\ultramodular}{\mathcal{M}}
\newcommand{\ultramodularL}[1][]{\ensuremath{\ultramodular^{L{#1}}}}
\newcommand{\ultramodularR}[1][]{\ensuremath{\ultramodular^{R{#1}}}}
\newcommand{\muB}{\ensuremath{\mu^{B}}}
\newcommand{\eventsC}{\ensuremath{\events_{C}}}

\begin{document}

@@

A principal result is that recasting the Kochen-Specker theorem in
terms of ``$\delta$-determinism,'' with~$\delta$ quantifying the
effect of finite-precision uncertainties on measurement outcomes,
reducing to the purely deterministic Kochen-Specker case for
$\delta=0$, can resolve the question of the effect of finite precision
and imperfections on contextuality.  A remarkable consequence of our
analysis is that not only does the quantum-mechanical requirement for
contextuality continue to hold, thus maintaining the Kochen-Specker
result for a \emph{non-vanishing} range of $\delta$, but that at a
\emph{certain fixed value} of $\delta$, there is a sharp transition,
parallel in its consequences to, say, a phase transition, to
\emph{non-contextuality}. %% ; that is, suddenly a quantum system can be
%% simulated by a non-contextual hidden variable theory.
%% , turning the
%% system effectively from a quantum system into a classical system. 
% that is, quantum mechanics suddenly fails to hold, and all variables
% can be specified simultaneously, turning the theory effectively into a
% classical theory that can be specified consistently by hidden
% variables.
% \yutsung{The part after ``that is''
%   looks strange.
% \begin{enumerate}
% \item The Kochen-Specker theorem is not about quantum mechanics, is about whether
% quantum mechanics can be simulated by a non-contextual deterministic hidden
% variable theory or not. Hence, ``quantum mechanics suddenly fails to hold''
% doesn't make a lot sense.
% \item The question is whether there exist a QIVPM means ``a classical theory
% consistently by hidden variables.'' At least, a QIVPM is not deterministic, and
% I need to check whether it qualities as a classical probabilistic hidden variable
% theory\ldots(?)
% \end{enumerate}
% }
These results provide a new insight into the Meyer-Mermin debate by
presenting a theory in which there is a \emph{parameter} interpolating
between what appeared previously to be irreconcilable interpretations
of contextuality.

@@

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections of
uncertain
events~\cite{Kolmogorov1950bibTeX,544199,Griffiths2003,Grabisch2016}. In the
quantum case, the events of interest are specified by \emph{projection
  operators} $P$ satisfying the condition $P^2=P$. These include the
empty projector~$\mathbb{0}$, the identity projector~$\mathbb{1}$,
projectors of the form $\proj{\phi}$ where $\ket{\phi}$ is a pure
quantum state (an element of a Hilbert space ${\cal H}$), sums of
\emph{orthogonal} projectors $P_0$ and~$P_1$ with $P_0P_1=\mathbb{0}$,
and products of \emph{commuting} projectors $P_0$ and~$P_1$ with
$P_0P_1=P_1P_0$. In a quantum probability
space~\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010,Abramsky2012},
each event~$P_{i}$ is mapped to a probability~$\mu(P_{i})$ using a
probability measure~$\mu:\events\rightarrow[0,1]$, where $\events$ is
the set of all events, projectors on a given Hilbert space, subject to the
following constraints: $\mu(\mathbb{0})=0$, $\mu(\mathbb{1})=1$,
$\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$, and for each pair
of \emph{orthogonal} projectors $P_{0}$ and $P_{1}$:
\begin{equation}
{\mu}\left(P_{0}+P_{1}\right)={\mu}\left(P_{0}\right)+{\mu}\left(P_{1}\right)\,.\label{eq:QuantumProbability-Addition}
\end{equation}
Given a Hilbert space ${\cal H}$ of dimension $d$ and a probability assignment
for every projector $P$, we can define the expectation value of an
observable~$\mathbf{O}$ having spectral decomposition
$\mathbf{O}=\sum_{i=1}^{d}\lambda_{i}P_{i}$, with eigenvalues $\lambda_i \in \mathbb{R}$, as:
\begin{equation}
\expval{\mathbf{O}}_{\mu}=\sum_{i=1}^{d}\lambda_{i}\mu(P_{i})\,.\label{eq:quantum-expectation}
\end{equation}

@@

First, for each interval $[\ell,r]$ we have the natural constraint
$0 \leq \ell \leq r \leq 1$ that guarantees that every element of the
interval can be interpreted as a conventional probability. We also
include $\imposs=[0,0]$ and $\necess=[1,1]$ as limiting intervals to
refer to the probability of impossible events (e.g., the null event)
and events that are certain (e.g., the event covering the entire
sample space).  For each interval $[\ell,r]$, we also need the dual
interval $[1-r,1-\ell]$ so that if one interval refers to the
probability of an event~$E$, the dual refers to the probability of the
event's complement $\overline{E}$.  For example, if we discover as a
result of an experiment that $\mu(E) = [0.2,0.3]$ for some event~$E$,
we may conclude that $\mu(\overline{E}) = [0.7,0.8]$ for the
complementary event $\overline{E}$. In addition to these simple
conditions, there are some subtle conditions on how intervals are
combined, which we discuss next.

@@

In the remainder of this section, we will show that our definition of
QIVPMs is consistent with classical IVPMs in the sense that a
restriction of QIVPMs to mutually commuting events recovers the definition of
classical IVPMs. To formalize this connection between QIVPMs and
classical IVPMs, we first note that in contrast to the case with
conventional quantum probability spaces, there is generally no unique
state~$\rho$ consistent with a QIVPM. We say that a QIVPM $\bar{\mu}$
is consistent with a state $\rho$ and projector~$P$ if the interval
$\bar{\mu}(P)$ contains the exact probability calculated by the Born
rule, i.e, if:
\begin{equation}
\muB_{\rho}\left(P\right)\in\bar{\mu}\left(P\right)
\end{equation}
where $\muB_{\rho}\left(P\right)=\Tr\left(\rho P\right)$
is the probability of the event~$P$ as determined by the Born
rule~\citep{Born1983bibTeX,Mermin2007,Jaeger2007}.

We generalize this consistency relation to sets of states and general
subspaces of events as follows. A set of events $\events'$ is a
\emph{subspace} of $\events$ if it contains $\mathbb{0}$
and~$\mathbb{1}$ and is closed under complements, sums, and products,
i.e., for any projector~$P\in\events'$, we have
$\mathbb{1}-P\in\events'$ and for each pair of commuting
projectors~$P_{0}\in\events'$ and $P_{1}\in\events'$, we have
$P_{0}+P_{1}-P_{0}P_{1}$ and $P_{0}P_{1}\in\events'$. Given a
probability measure $\bar{\mu}$ and a subspace of events $\events'$,
we denote the collection of all states that are consistent with
$\bar{\mu}$ on every projector in $\events'$ as
$\coreBorn(\bar{\mu},\events')$:
\begin{equation}
\label{eq:hbar}
\coreBorn\left(\bar{\mu},\events'\right)=\set{\rho}{\forall P\in \events',~\muB_{\rho}\left(P\right)\in\bar{\mu}\left(P\right)}.
\end{equation}

The proof that QIVPMs become classical IVPMs when the space of events
includes only mutually commuting events is presented in detail in the
long version of this paper~\cite{TaiThesis2018}. The idea of the
proof is as follows. Given a subspace of mutually commuting
events~$\eventsC$, we can construct a partial orthonormal
basis by diagonalization and complete it to a full orthonormal basis
$\overline{\eventsC}$. We then build a bijection between
the QIVPM on the set of projectors associated with this basis and the
set of classical events corresponding to this basis. This
correspondence leads to a crucial property in our setting.

@@

This theorem follows directly from our
correspondence~\cite{TaiThesis2018} and Shapley's
theorem~\cite{Shapley1971,GilboaSchmeidler1994,NgMoYeh1997,Grabisch2016},
which proves that a classical IVPM always contains at least one state $\rho$
that is consistent with every event in $\eventsC$. In contrast, as we
illustrate in Sec.~\ref{sec:Gleason}, the set of states
$\coreBorn\left(\bar{\mu},\events'\right)$ that is consistent with
some arbitrary (not necessarily commuting) events $\events'$ and a
particular QIVPM $\bar{\mu}$ may be \emph{empty}.

For example, a well-defined  
QIVPM, assuming the three intervals $\imposs$, $\necess$, and
$\unknown=\left[0,1\right]$, is:
\begin{equation}
\bar{\mu}_{1}\left(P\right)=\begin{cases}
\imposs\,, & \text{if }P=\mathbb{0}\text{ ;}\\
\necess\,, & \text{if }P=\mathbb{1}\text{ ;}\\
\unknown\,, & \text{otherwise.}
\end{cases}\label{eq:no-information-IVPM}
\end{equation}
This measure represents the beliefs of an experimenter with no prior
knowledge about the particular quantum system in question; any quantum
state would be consistent with $\bar{\mu}_{1}$ on any projectors. That
is, $\coreBorn\left(\bar{\mu}_{1},\events\right)$ is the set of all
states. The subspace
$\events_{0}=\left\{ \mathbb{0},\proj{0},\proj{1},\mathbb{1}\right\} $
is a subspace of commuting events. Restricting $\bar{\mu}_{1}$
to $\events_{0}$, $\bar{\mu}_{1}|_{\events_{0}}$, gives a measure
$\bar{\mu}_{2}$:
\begin{equation}
\begin{aligned}\bar{\mu}_{2}(\mathbb{0}) & =\imposs & \bar{\mu}_{2}(\proj{0}) & =\unknown\\
\bar{\mu}_{2}(\mathbb{1}) & =\necess & \bar{\mu}_{2}(\proj{1}) & =\unknown
\end{aligned}
\label{eq:classical-no-information-IVPM}
\end{equation}
that corresponds to a classical IVPM. Indeed, if we identify the
states~$\proj{0}$ and $\proj{1}$ with the head and the tail of a coin,
respectively, then $\bar{\mu}_{2}$ represents the situation in which
an experimenter knows nothing about a coin before she tosses
it. Moreover, for each $p\in\left[0,1\right]$, the
state~$p\proj{0}+\left(1-p\right)\proj{1}$ represents a classical coin
with probability~$p$ to get heads, and
$p\proj{0}+\left(1-p\right)\proj{1}$ is consistent
with~$\bar{\mu}_{2}$ on $\events_{0}$ because any classical coin
should be consistent with the situation for which an experimenter has
no prior knowledge.

@@

The expectation value of an observable $\mathbf{O}$ in a Hilbert space
${\cal H}$ of dimension $d$ relative to a 0-deterministic QIVPM is fully
determinate and is equal to one of the eigenvalues $\lambda_i$ of that
observable. To see this, note that given an orthonormal basis
$\Omega=\{ \ket{\psi_0},\ket{\psi_1},\ldots,\ket{\psi_{d-1}}\}$, a
0-deterministic QIVPM must map exactly one of the projectors
$\proj{\psi_i}$ to $\necess$ and all others to $\imposs$. This is
because, by Eq.~(\ref{eq:necess}), we have
$\bar{\mu}(\Sigma_{j=0}^{d-1}{\proj{\psi_j}})=\necess$ and by inductively
applying Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}),
we must have one of the $\bar{\mu}(\proj{\psi_i})=\necess$ and all
others mapped to $\imposs$. Given any state~$\rho$ that is consistent
with this QIVPM on all the projectors in $\Omega$, we have by
Eq.~(\ref{eq:hbar}) that $\muB_{\rho}$ must also map
exactly one of the projectors in $\Omega$ to 1 and all others to 0. If
an observable has a spectral decomposition along $\Omega$ then, by
Eq.~(\ref{eq:quantum-expectation}), its expectation value relative to
$\muB_{\rho}$ is the eigenvalue $\lambda_i$ whose projector
is mapped to 1. It therefore follows, by
Eq.~(\ref{eq:quantum-interval-expectation-core}), that the expectation
value relative to the 0-deterministic~${\bar{\mu}}$ is fully
determinate and lies in the interval $[\lambda_i,\lambda_i]$.

@@

\noindent The observables are constructed using the Pauli matrices
$\left\{ \mathbb{1},\sigma_{x},\sigma_{y},\sigma_{z}\right\}$ whose
eigenvalues are all either $1$ or $-1$
\cite{Redhead1987-REDINA,544199,Griffiths2003,Jaeger2007,Mermin2007}.
They are arranged such that in
each row and column, \emph{except the column $j=2$}, every observable
is the product of the other two. In the $j=2$ column, we have instead
that
$\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$. Now
assume a 0-deterministic QIVPM~${\bar{\mu}}$; the expectation values
of the observables in each row relative to this 0-deterministic QIVPM
are fully determinate and must lie in either the interval $[1,1]$ or
the interval $[-1,-1]$ depending on which eigenvalue is the one whose
associated projector is certain. Since the product of any two
observables in a row is equal to the third, there must be an even
number of occurrences of the interval $[-1,-1]$ in each row and hence
in the entire table~\citep{TaiThesis2018}. However, looking at the expectation values of the
observables in each column, there must be an even number of
occurrences of the interval $[-1,-1]$ in the first two columns and an
odd number in the $j=2$ column and hence in the entire table~\citep{TaiThesis2018}. The
contradiction implies the non-existence of the assumed 0-deterministic
QIVPM. 
%% We therefore prove a stronger statement of contextuality that 
%% includes the effects of finite-precision.

@@

As is the case for conventional, infinitely-precise, quantum
probability measures, the theorem is only applicable to dimensions
$d \geq 3$. Indeed when the Hilbert space has dimension 2, it is
straightforward to construct a 0-deterministic QIVPM as
follows. Consider a non-contextual hidden variable model (e.g., as
proposed by Bell or
Kochen-Specker~\cite{BELL_1966,kochenspecker1967}). Such a model
assigns definite values to all observables at all times, and hence
assigns a \emph{determinate} probability (0 or 1) to each event. Such
a probability measure directly induces a 0-deterministic QIVPM by
changing 0 to $\imposs$ and 1 to $\necess$.

>>>>

\documentclass[english,reprint, aps, prl,superscriptaddress, showpacs,
showkeys, longbibliography, amsmath, amssymb, floatfix]{revtex4-1}

\usepackage{fullpage}
\usepackage{cmap}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[french,main=english]{babel}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{bbold}
\usepackage{wesa}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage[backref=false]{hyperref}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[braket,qm]{qcircuit}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{framed}

@@

\DeclareUnicodeCharacter{0229}{\c{e}}

\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\qevents}{\ensuremath{\mathcal{E}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\imposs}{{\text{\wesa{impossible}}}}
\newcommand{\likely}{{\text{\wesa{likely}}}}
\newcommand{\unlikely}{{\text{\wesa{unlikely}}}}
\newcommand{\necess}{{\text{\wesa{certain}}}}
\newcommand{\unknown}{{\text{\wesa{unknown}}}}
\newcommand{\midd}{{\text{\wesa{middle}}}}
\newcommand{\fket}[1]{{|#1\rangle}}
\newcommand{\fproj}[1]{|#1\rangle\langle #1|}
\newcommand{\proj}[1]{\op{#1}{#1}}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\allowdisplaybreaks
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{L{#1}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{R{#1}}}}
\setcounter{secnumdepth}{3}
% https://tex.stackexchange.com/questions/345284/can-i-have-an-almost-non-breaking-space-in-latex
% Change non-breaking space to \nolinebreak[1] 
% Hence, it will not hyphenate too much surrounding words.
\newcommand{\rmd}{d}
\newcommand{\rmi}{i}
\lineskiplimit=1pt
\newcommand{\ultramodular}{\mathcal{M}}
\newcommand{\ultramodularL}[1][]{\ensuremath{\ultramodular^{L{#1}}}}
\newcommand{\ultramodularR}[1][]{\ensuremath{\ultramodular^{R{#1}}}}
\newcommand{\muB}{\ensuremath{\mu^{B}}}
\newcommand{\eventsC}{\ensuremath{\events_{C}}}

\newcommand{\amr}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{green}{Amr says: #1}\end{minipage}\end{framed}}
\newcommand{\yutsung}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}\end{framed}}
\newcommand{\gerardo}[1]{\fbox{\begin{minipage}{0.9\linewidth}\color{OliveGreen}{Gerardo says: #1}\end{minipage}}}
\newcommand{\andy}[1]{\fbox{\begin{minipage}{0.9\linewidth}\color{blue}{Andy says: #1}\end{minipage}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

@@

We begin with the concept of $\delta$-determinism where~$\delta$
quantifies the effect of finite-precision uncertainties on measurement
outcomes. We recast the Kochen-Specker theorem using
$\delta$-determinism to quantify the effects of finite precision and
imperfections on contextuality. When $\delta=0$, the generalized
theorem reduces to the conventional one but as $\delta$ varies from 0
to 1 we note a transition from contextuality to non-contextuality.
% A principal result is that recasting the Kochen-Specker theorem in 
% terms of ``$\delta$-determinism,'' with~$\delta$ quantifying the 
% effect of finite-precision uncertainties on measurement outcomes, 
% reducing to the purely deterministic Kochen-Specker case for 
% $\delta=0$, can resolve the question of the effect of finite precision 
% and imperfections on contextuality.  
In more detail, the quantum-mechanical requirement for contextuality
continues to hold, thus maintaining the Kochen-Specker result for a
\emph{non-vanishing} range of $\delta$, but that at a \emph{certain
  fixed value} of $\delta$, there is a sharp transition, parallel in
its consequences to, say, a phase transition, to
\emph{non-contextuality}. %% ; that is, suddenly a quantum system can be
%% simulated by a non-contextual hidden variable theory.
%% , turning the
%% system effectively from a quantum system into a classical system. 
% that is, quantum mechanics suddenly fails to hold, and all variables
% can be specified simultaneously, turning the theory effectively into a
% classical theory that can be specified consistently by hidden
% variables.
% \yutsung{The part after ``that is''
%   looks strange.
% \begin{enumerate}
% \item The Kochen-Specker theorem is not about quantum mechanics, is about whether
% quantum mechanics can be simulated by a non-contextual deterministic hidden
% variable theory or not. Hence, ``quantum mechanics suddenly fails to hold''
% doesn't make a lot sense.
% \item The question is whether there exist a QIVPM means ``a classical theory
% consistently by hidden variables.'' At least, a QIVPM is not deterministic, and
% I need to check whether it qualities as a classical probabilistic hidden variable
% theory\ldots(?)
% \end{enumerate}
% }
These results provide a new insight into the Meyer-Mermin debate by
presenting a theory in which there is a \emph{parameter} interpolating
between what appeared previously to be irreconcilable interpretations
of contextuality.

@@

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections of
uncertain
events~\cite{Kolmogorov1950bibTeX,544199,Griffiths2003,Grabisch2016}. In the
quantum case, the events of interest are specified by \emph{projection
  operators} $P$ satisfying the condition $P^2=P$. These include the
empty projector~$\mathbb{0}$, the identity projector~$\mathbb{1}$,
projectors of the form $\proj{\phi}$ where $\ket{\phi}$ is a pure
quantum state (an element of a Hilbert space ${\cal H}$), sums of
\emph{orthogonal} projectors~$P_0$ and~$P_1$ with $P_0P_1=\mathbb{0}$,
and products of \emph{commuting} projectors $P_0$ and~$P_1$ with
$P_0P_1=P_1P_0$. In a quantum probability
space~\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010,Abramsky2012},
each event~$P_{i}$ is mapped to a probability~$\mu(P_{i})$ using a
probability measure~$\mu:\events\rightarrow[0,1]$, where $\events$ is
the set of all events, projectors on a given Hilbert space, subject to the
following constraints: $\mu(\mathbb{0})=0$, $\mu(\mathbb{1})=1$,
$\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$, and for each pair
of \emph{orthogonal} projectors $P_{0}$ and $P_{1}$:
\begin{equation}
{\mu}\left(P_{0}+P_{1}\right)={\mu}\left(P_{0}\right)+{\mu}\left(P_{1}\right)\,.\label{eq:QuantumProbability-Addition}
\end{equation}
Given a Hilbert space ${\cal H}$ of dimension $d$ and a probability assignment
for every projector $P$, we can define the expectation value of an
observable~$\mathbf{O}$ having spectral decomposition
$\mathbf{O}=\sum_{i=1}^{d}\lambda_{i}P_{i}$, with eigenvalues $\lambda_i \in \mathbb{R}$, as:
\begin{equation}
\expval{\mathbf{O}}_{\mu}=\sum_{i=1}^{d}\lambda_{i}\mu(P_{i})\,.\label{eq:quantum-expectation}
\end{equation}

@@

First, for each interval $[\ell,r]$ we have the natural constraint
$0 \leq \ell \leq r \leq 1$ that guarantees that every element of the
interval can be interpreted as a conventional probability. We also
include $\imposs=[0,0]$ and $\necess=[1,1]$ as limiting intervals to
refer to the probability of impossible events (e.g., the null event)
and events that are certain (e.g., the event covering the entire
sample space).  For each interval $[\ell,r]$, we also need the dual
interval $[1-r,1-\ell]$ so that if one interval refers to the
probability of an event~$E$, the dual refers to the probability of the
event's complement $\overline{E}$.  For example, if we discover as a
result of an experiment that $\mu(E) = [0.2,0.3]$ for some event~$E$,
we may conclude that $\mu(\overline{E}) = [0.7,0.8]$ for the
complementary event~$\overline{E}$. In addition to these simple
conditions, there are some subtle conditions on how intervals are
combined, which we discuss next.

@@

We now show that our definition of QIVPMs is consistent with classical
IVPMs in the sense that a restriction of QIVPMs to mutually commuting
events recovers the definition of classical IVPMs. To formalize this
connection, we first note that in contrast to the case with
conventional quantum probability spaces, there is generally no unique
state~$\rho$ consistent with a QIVPM. We say that a QIVPM $\bar{\mu}$
is \emph{consistent} with a state $\rho$ and projector~$P$ if the
interval $\bar{\mu}(P)$ contains the exact probability calculated by
the Born rule, i.e, if:
\begin{equation}
\muB_{\rho}\left(P\right)\in\bar{\mu}\left(P\right)
\end{equation}
where $\muB_{\rho}\left(P\right)=\Tr\left(\rho P\right)$
is the probability of the event~$P$ as determined by the Born
rule~\citep{Born1983bibTeX,Mermin2007,Jaeger2007}.

We generalize this consistency relation to sets of states and general
subspaces of events as follows. A set of events $\events'$ is a
\emph{subspace} of $\events$ if it contains $\mathbb{0}$
and~$\mathbb{1}$ and is closed under complements, sums, and products,
i.e., for any projector~$P\in\events'$, we have
$\mathbb{1}-P\in\events'$ and for each pair of commuting
projectors~$P_{0}\in\events'$ and $P_{1}\in\events'$, we have
$P_{0}+P_{1}-P_{0}P_{1}$ and $P_{0}P_{1}\in\events'$. Given a
probability measure $\bar{\mu}$ and a subspace of events $\events'$,
we define the \emph{core} of $\bar{\mu}$ relative to $\events'$ as the
collection of all states $\rho$ that are consistent with $\bar{\mu}$
on every projector in $\events'$ as $\coreBorn(\bar{\mu},\events')$:
\begin{equation}
\label{eq:hbar}
\coreBorn\left(\bar{\mu},\events'\right)=\set{\rho}{\forall P\in \events',~\muB_{\rho}\left(P\right)\in\bar{\mu}\left(P\right)}.
\end{equation}

The proof that QIVPMs become classical IVPMs when the space of events
includes only mutually commuting events is presented in detail in a
forthcoming thesis~\cite{TaiThesis2018}. The idea of the proof is as
follows. Given a subspace of mutually commuting events~$\eventsC$, we
can construct a partial orthonormal basis by diagonalization and
complete it to a full orthonormal basis $\overline{\eventsC}$. We then
build a bijection between the QIVPM on the set of projectors
associated with this basis and the set of classical events
corresponding to this basis. Using this correspondence together with a
classical result by
Shapley~\cite{Shapley1971,GilboaSchmeidler1994,NgMoYeh1997,Grabisch2016},
we can establish that for the special case of commuting events, a
QIVPM will always have a non-empty core~\footnote{Shapley's result
  proves that a classical IVPM always contains at least one state that
  is consistent with every event.}.

@@

The importance of this property is that it is violated in
finite-precision versions of Gleason's theorem
(cf. Sec.~\ref{sec:Gleason}). As we will show, the set of states
$\coreBorn\left(\bar{\mu},\events'\right)$ that is consistent with
some arbitrary (not necessarily commuting) events~$\events'$ and a
particular QIVPM $\bar{\mu}$ may be \emph{empty}.

% For example, a well-defined QIVPM, assuming the three intervals
% $\imposs$, $\necess$, and $\unknown=\left[0,1\right]$, is:
% \begin{equation}
% \bar{\mu}_{1}\left(P\right)=\begin{cases}
% \imposs\,, & \text{if }P=\mathbb{0}\text{ ;}\\
% \necess\,, & \text{if }P=\mathbb{1}\text{ ;}\\
% \unknown\,, & \text{otherwise.}
% \end{cases}\label{eq:no-information-IVPM}
% \end{equation}
% This measure represents the beliefs of an experimenter with no prior
% knowledge about the particular quantum system in question; any quantum
% state would be consistent with $\bar{\mu}_{1}$ on any projectors. That
% is, $\coreBorn\left(\bar{\mu}_{1},\events\right)$ is the set of all
% states. The subspace
% $\events_{0}=\left\{ \mathbb{0},\proj{0},\proj{1},\mathbb{1}\right\} $
% is a subspace of commuting events. Restricting $\bar{\mu}_{1}$
% to $\events_{0}$, $\bar{\mu}_{1}|_{\events_{0}}$, gives a measure
% $\bar{\mu}_{2}$:
% \begin{equation}
% \begin{aligned}\bar{\mu}_{2}(\mathbb{0}) & =\imposs & \bar{\mu}_{2}(\proj{0}) & =\unknown\\
% \bar{\mu}_{2}(\mathbb{1}) & =\necess & \bar{\mu}_{2}(\proj{1}) & =\unknown
% \end{aligned}
% \label{eq:classical-no-information-IVPM}
% \end{equation}
% that corresponds to a classical IVPM. Indeed, if we identify the
% states~$\proj{0}$ and $\proj{1}$ with the head and the tail of a coin,
% respectively, then $\bar{\mu}_{2}$ represents the situation in which
% an experimenter knows nothing about a coin before she tosses
% it. Moreover, for each $p\in\left[0,1\right]$, the
% state~$p\proj{0}+\left(1-p\right)\proj{1}$ represents a classical coin
% with probability~$p$ to get heads, and
% $p\proj{0}+\left(1-p\right)\proj{1}$ is consistent
% with~$\bar{\mu}_{2}$ on $\events_{0}$ because any classical coin
% should be consistent with the situation for which an experimenter has
% no prior knowledge.

@@

The expectation value of an observable $\mathbf{O}$ in a Hilbert space
${\cal H}$ of dimension $d$ relative to a 0-deterministic QIVPM is
fully determinate and is equal to one of the eigenvalues $\lambda_i$
of that observable. To see this, note that given an orthonormal basis
$\Omega=\{ \ket{\psi_0},\ket{\psi_1},\ldots,\ket{\psi_{d-1}}\}$, a
0-deterministic QIVPM must map exactly one of the projectors
$\proj{\psi_i}$ to $\necess$ and all others to $\imposs$. This is
because, by Eq.~(\ref{eq:necess}), we have
$\bar{\mu}(\Sigma_{j=0}^{d-1}{\proj{\psi_j}})=\necess$ and by
inductively applying
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}), we must
have one of the $\bar{\mu}(\proj{\psi_i})=\necess$ and all others
mapped to $\imposs$. Given any state~$\rho$ that is consistent with
this QIVPM on all the projectors in $\Omega$, we have by
Eq.~(\ref{eq:hbar}) that $\muB_{\rho}$ must also map exactly one of
the projectors in $\Omega$ to 1 and all others to 0. If an observable
has a spectral decomposition along~$\Omega$ then, by
Eq.~(\ref{eq:quantum-expectation}), its expectation value relative
to~$\muB_{\rho}$ is the eigenvalue $\lambda_i$ whose projector is
mapped to 1. It therefore follows, by
Eq.~(\ref{eq:quantum-interval-expectation-core}), that the expectation
value relative to the 0-deterministic~${\bar{\mu}}$ is fully
determinate and lies in the interval $[\lambda_i,\lambda_i]$.

@@

\noindent The observables are constructed using the Pauli matrices
$\left\{ \mathbb{1},\sigma_{x},\sigma_{y},\sigma_{z}\right\}$ whose
eigenvalues are all either~$1$ or $-1$
\cite{Redhead1987-REDINA,544199,Griffiths2003,Jaeger2007,Mermin2007}.
They are arranged such that in
each row and column, \emph{except the column $j=2$}, every observable
is the product of the other two. In the $j=2$ column, we have instead
that
$\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$. Now
assume a 0-deterministic QIVPM~${\bar{\mu}}$; the expectation values
of the observables in each row relative to this 0-deterministic QIVPM
are fully determinate and must lie in either the interval $[1,1]$ or
the interval $[-1,-1]$ depending on which eigenvalue is the one whose
associated projector is certain. Since the product of any two
observables in a row is equal to the third, there must be an even
number of occurrences of the interval $[-1,-1]$ in each row and hence
in the entire table~\citep{TaiThesis2018}. However, looking at the expectation values of the
observables in each column, there must be an even number of
occurrences of the interval $[-1,-1]$ in the first two columns and an
odd number in the $j=2$ column and hence in the entire table~\citep{TaiThesis2018}. The
contradiction implies the non-existence of the assumed 0-deterministic
QIVPM. 
%% We therefore prove a stronger statement of contextuality that 
%% includes the effects of finite-precision.

@@

As is the case for conventional, infinitely-precise, quantum
probability measures, the theorem is only applicable to dimensions
$d \geq 3$. Indeed when the Hilbert space has dimension 2, it is
straightforward to construct a 0-deterministic QIVPM as
follows. Consider a non-contextual hidden variable model for $d=2$
(e.g., as proposed by Bell or
Kochen-Specker~\cite{BELL_1966,kochenspecker1967}). Such a
two-dimensional model assigns definite values to all observables at
all times, and hence assigns a \emph{determinate} probability (0 or 1)
to each event. This probability measure directly induces a
0-deterministic QIVPM by changing 0 to $\imposs$ and 1 to $\necess$,
and therefore a $\delta$-deterministic one.
